---
title: "DeSimone_Assignment 2"
author: "Heather DeSimone"
date: "2/19/2022"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
##First I have loaded in my data frame and called a summary of the information.
```{r}
DF=read.csv("C:/Users/hdesi/Desktop/MBA/Machine Learning/UniversalBank.csv")
library(dplyr)
DF <- DF %>% relocate(Personal.Loan, .after = CreditCard)
summary(DF)
```
##Next I will remove the 2 variables that will not be used in my classification/prediction: ID and Zip Code.
##I have also converted a few attributes over to factors - these attributes classify a yes (1) or no (0) response.I have called a summary to check my work.
```{r}
DF$ID<-NULL
DF$ZIP.Code<-NULL
DF$Personal.Loan=as.factor(DF$Personal.Loan)
DF$Securities.Account=as.factor(DF$Securities.Account)
DF$CD.Account=as.factor(DF$CD.Account)
DF$Online=as.factor(DF$Online)
DF$CreditCard=as.factor(DF$CreditCard)
summary(DF)
```
##I will now load the caret and class libraries.
```{r}
library(caret)
library(class)
```
##Next I have created a new data set for the normalization process - I have removed the target variable: Personal Loan as we cannot normalize it.
##I have also removed the attributes that were factored
```{r}
Normalization_DF <- data.frame(DF)
Normalization_DF$Personal.Loan<-NULL
Normalization_DF$Securities.Account<-NULL
Normalization_DF$CD.Account<-NULL
Normalization_DF$Online<-NULL
Normalization_DF$CreditCard<-NULL
summary(Normalization_DF)
```
##I will now normalize the data.
```{r}
Norm_model <- preProcess(Normalization_DF, 
                         method = c("center", "scale"))
loan_norm=predict(Norm_model,Normalization_DF)
summary(loan_norm)
```
##I will now add the attributes back in that I removed for normalization.
```{r}
loan_norm$Personal.Loan=DF$Personal.Loan
loan_norm$Securities.Account=DF$Securities.Account
loan_norm$CD.Account=DF$CD.Account
loan_norm$Online=DF$Online
loan_norm$CreditCard=DF$CreditCard
summary(loan_norm)
```
##I will now separate my data into training and validating sets - training = 60% and validation = 40%.
```{r}
Train_Index = createDataPartition(DF$Personal.Loan,p=0.6, list=FALSE)
Train.df=loan_norm[Train_Index,]
Validation.df=loan_norm[-Train_Index,] 
```
##Question #1
##I will now input the attributes of the 1st customer for prediction.
```{r}
To_Predict=data.frame(Age=40, Experience=10,
                      Income=84,Family=2,
                      CCAvg=2,Education=2,
                      Mortgage=0,
                      Securities.Account=0,
                      CD.Account=0,
                      Online=1,
                      CreditCard=1)

print(To_Predict)
```
##I will remove the attributes that were factored.
```{r}
To_Predict_norm=To_Predict
To_Predict_norm$Personal.Loan<-NULL
To_Predict_norm$Securities.Account<-NULL
To_Predict_norm$CD.Account<-NULL
To_Predict_norm$Online<-NULL
To_Predict_norm$CreditCard<-NULL
```
##I will now normalize the data.
```{r}
To_Predict_norm=predict(Norm_model,To_Predict_norm)
```
##I will now add the attributes back in that I removed for normalization.
```{r}
To_Predict_norm$Personal.Loan<-To_Predict$Personal.Loan
To_Predict_norm$Securities.Account<-To_Predict$Securities.Account
To_Predict_norm$CD.Account<-To_Predict$CD.Account
To_Predict_norm$Online<-To_Predict$Online
To_Predict_norm$CreditCard<-To_Predict$CreditCard
print(To_Predict_norm)
```
##I will now use the knn function to make my prediction.
```{r}
Train.df <- Train.df %>% relocate(Personal.Loan, .after = CreditCard)
Prediction <-knn(train=Train.df[1:11], 
                 test=To_Predict_norm[1:11],
                 cl=Train.df$Personal.Loan,
                 k=1)
print(Prediction)
```
##This customer is predicted NOT to accept the personal loan

##Question #2
##I will now build the knn model that will give the best value of k that balances between overfitting and underfitting.
```{r}
set.seed(123)

fitControl <- trainControl(method = "repeatedcv",
                           number = 3,
                           repeats = 2)

searchGrid=expand.grid(k = 1:10)

Knn.model=train(Personal.Loan~., 
                data=Train.df,
                method='knn',
                tuneGrid=searchGrid,
                trControl = fitControl,)

Knn.model
```
##The best k value to use is 3.
##Question #3
##First I will use the predict function of the caret package.
```{r}
predictions<-predict(Knn.model,Validation.df)
```
#Now I will compute the confusion matrix using the caret package.
```{r}
confusionMatrix(predictions,Validation.df$Personal.Loan)

```

##Question #4
##I will now input the attributes of the 2nd customer for prediction. *Note that this customer information is the same as the 1st customer.
```{r}
To_Predict2=data.frame(Age=40, Experience=10,
                      Income=84,Family=2,
                      CCAvg=2,Education=2,
                      Mortgage=0,
                      Securities.Account=0,
                      CD.Account=0,
                      Online=1,
                      CreditCard=1)

print(To_Predict2)
```
##I will remove the attributes that were factored.
```{r}
To_Predict_norm2=To_Predict2
To_Predict_norm2$Personal.Loan<-NULL
To_Predict_norm2$Securities.Account<-NULL
To_Predict_norm2$CD.Account<-NULL
To_Predict_norm2$Online<-NULL
To_Predict_norm2$CreditCard<-NULL
```
##I will now normalize the data.
```{r}
To_Predict_norm2=predict(Norm_model,To_Predict_norm2)
```
##I will now add the attributes back in that I removed for normalization.
```{r}
To_Predict_norm2$Personal.Loan<-To_Predict2$Personal.Loan
To_Predict_norm2$Securities.Account<-To_Predict2$Securities.Account
To_Predict_norm2$CD.Account<-To_Predict2$CD.Account
To_Predict_norm2$Online<-To_Predict2$Online
To_Predict_norm2$CreditCard<-To_Predict2$CreditCard
print(To_Predict_norm2)
```
##I will now use the knn function to make my prediction.I am using k=3 as it is the best k value.
```{r}
Train.df <- Train.df %>% relocate(Personal.Loan, .after = CreditCard)
Prediction <-knn(train=Train.df[1:11], 
                 test=To_Predict_norm2[1:11],
                 cl=Train.df$Personal.Loan,
                 k=3)
print(Prediction)
```

##This customer is predicted NOT to take out the personal loan
##Question 5
##I will now repartition my data into training (50%), Validation (30%), and test (20%).
```{r}
Train_Index2 = createDataPartition(DF$Personal.Loan,p=0.5, list=FALSE)
Train.df2=loan_norm[Train_Index2,]
Validation.df2=loan_norm[-Train_Index2,] 

```

##I will now input the attributes of the 1st customer for prediction.
```{r}
To_Predict3=data.frame(Age=40, Experience=10,
                      Income=84,Family=2,
                      CCAvg=2,Education=2,
                      Mortgage=0,
                      Securities.Account=0,
                      CD.Account=0,
                      Online=1,
                      CreditCard=1)

print(To_Predict3)
```
##I will remove the attributes that were factored.
```{r}
To_Predict_norm3=To_Predict3
To_Predict_norm3$Personal.Loan<-NULL
To_Predict_norm3$Securities.Account<-NULL
To_Predict_norm3$CD.Account<-NULL
To_Predict_norm3$Online<-NULL
To_Predict_norm3$CreditCard<-NULL
```
##I will now normalize the data.
```{r}
To_Predict_norm3=predict(Norm_model,To_Predict_norm3)
```
##I will now add the attributes back in that I removed for normalization.
```{r}
To_Predict_norm3$Personal.Loan<-To_Predict3$Personal.Loan
To_Predict_norm3$Securities.Account<-To_Predict3$Securities.Account
To_Predict_norm3$CD.Account<-To_Predict3$CD.Account
To_Predict_norm3$Online<-To_Predict3$Online
To_Predict_norm3$CreditCard<-To_Predict3$CreditCard
print(To_Predict_norm3)
```
##I will now use the knn function to make my prediction.
```{r}
Train.df2 <- Train.df2 %>% relocate(Personal.Loan, .after = CreditCard)
Prediction <-knn(train=Train.df2[1:11], 
                 test=To_Predict_norm2[1:11],
                 cl=Train.df2$Personal.Loan,
                 k=3)
print(Prediction)
```
##Now I will create my confusion matrix
##First I will use the predict function of the caret package.
```{r}
predictions2<-predict(Knn.model,Validation.df2)
```

#Now I will compute the confusion matrix using the caret package.
```{r}
confusionMatrix(predictions2,Validation.df2$Personal.Loan)

```
##The prediction stands that the customer will not take out the personal loan. Accuracy from the prior matrix to this one has increased.
