---
title: "DeSimone_MS64060_Assignment 4"
author: "Heather DeSimone"
date: "3/19/2022"
output:
  word_document: default
  pdf_document: default
  html_document: default
---


##First I have loaded in my data frame and called a summary of the information.
```{r}
df.original=read.csv("C:/Users/hdesi/Desktop/MBA/Machine Learning/Pharmaceuticals.csv")
df.original.names.numerical=read.csv("C:/Users/hdesi/Desktop/MBA/Machine Learning/Pharmaceuticals.csv")[,c(1:11)]
df=read.csv("C:/Users/hdesi/Desktop/MBA/Machine Learning/Pharmaceuticals.csv")[,c(3:11)]
library(ISLR)
library(dplyr)
library(tidyverse)
library(factoextra)
set.seed(123)
summary(df)
```

```{r}
df <- scale(df) #z-score
distance <- get_dist(df)
fviz_dist(distance)
```
##Will try k=4 first since 4 is the median distance shown in the above graph
##Will first use 25 restarts as it seems to be a typical number of random centroids to start with (based on the internet community)
```{r}
k4 <- kmeans(df, centers = 4, nstart = 25)
k4$centers
k4$size
## 4, 8, 6, 3

##21 data points total so lets look at where the 1st, last, and middle data points are 
k4$cluster[1] ##cluster 2
k4$cluster[10] ##cluster 2
k4$cluster[21] ##cluster 2

fviz_cluster(k4, data = df) ##Visual
```
##Lets see what happens with k=5
```{r}
k5 <- kmeans(df, centers = 5, nstart = 25)
k5$centers
k5$size
## 4, 2, 3, 8, 4

##21 data points total so lets look at where the 1st, last, and middle data points are 
k5$cluster[1] ##cluster 4
k5$cluster[10] ##cluster 4
k5$cluster[21] ##cluster 4
##These data points are all falling in the same bucket - close together

fviz_cluster(k5, data = df) ##Visual
```

##Lets see what happens with k=3
```{r}
k3 <- kmeans(df, centers = 3, nstart = 25)
k3$centers
k3$size
## 11, 4, 6

##21 data points total so lets look at where the 1st, last, and middle data points are 
k3$cluster[1] ##cluster 1
k3$cluster[10] ##cluster 1
k3$cluster[21] ##cluster 1
##These data points are all falling in the same bucket - close together

fviz_cluster(k3, data = df) ##Visual
```
##So far it looks like k=4 is the best option as it disperses the data somewhat evenly without splitting the data too drastically (a cluster of 1 or 2 observations doesn't seem to be helpful)

```{r}
fviz_nbclust(df, kmeans, method = "wss")
```
##Based on the elbow method graph, it looks like K = 4, 5, or 6 is the optimal number of clusters
```{r}
fviz_nbclust(df, kmeans, method = "silhouette")
```
##Based on the silhouette method, K=5 is the optimal number of clusters. We will use k=5
##I will use k=5 and now use the manhattan distance for clustering the data
```{r}
library(flexclust)
set.seed(123)
k5.manhattan = kcca(df, k=5, kccaFamily("kmedians"))
k5
```
##Based on our two models, it seems that the firms in cluster 5 are the top performers.


##Now I will examine 2 specific attributes - ROE & ROA - The higher the ROE & ROA are, the better the firm is performing and will probably continue to perform well.
```{r}
ROE.ROA.DF=read.csv("C:/Users/hdesi/Desktop/MBA/Machine Learning/Pharmaceuticals.csv")[,c(6,7)]
set.seed(123)
summary(ROE.ROA.DF)
```
##Looking for the best k value
```{r}
ROE.ROA.DF <- scale(ROE.ROA.DF) #z-score
distance <- get_dist(ROE.ROA.DF)
fviz_dist(distance)

fviz_nbclust(ROE.ROA.DF, kmeans, method = "wss")##elbow
fviz_nbclust(ROE.ROA.DF, kmeans, method = "silhouette")

```
##Based on both methods, k=2 or k=3 is optimal - we will use 3 as 2 is too insignificant
```{r}
ROA.ROE.PERFORMANCE <- kmeans(ROE.ROA.DF, centers = 3, nstart = 25)
ROA.ROE.PERFORMANCE$centers
ROA.ROE.PERFORMANCE$size
## 3, 8, 10

fviz_cluster(ROA.ROE.PERFORMANCE, data = ROE.ROA.DF) ##Visual
```
##Firms in cluster 1 have the highest ROA and ROE - data points 11,17, & 21
##Best: 11 - GlaxoSmithKline plc
##Also High Performers:    17 - Pfizer Inc    21 - Wyeth


##We will examine the market capitalization and the price to earnings ratio to determine the worth of a firm (in terms of investing).
```{r}
FIRM.WORTH.DF=read.csv("C:/Users/hdesi/Desktop/MBA/Machine Learning/Pharmaceuticals.csv")[,c(3,5)]
set.seed(123)
summary(FIRM.WORTH.DF)
```
##Looking for the best k value
```{r}
FIRM.WORTH.DF <- scale(FIRM.WORTH.DF) #z-score
distance <- get_dist(FIRM.WORTH.DF)
fviz_dist(distance)

fviz_nbclust(FIRM.WORTH.DF, kmeans, method = "wss")##elbow
fviz_nbclust(FIRM.WORTH.DF, kmeans, method = "silhouette")

```
##Based on both methods, k=3 is optimal 
```{r}
FIRM.WORTH <- kmeans(FIRM.WORTH.DF, centers = 3, nstart = 25)
FIRM.WORTH$centers
FIRM.WORTH$size
## 5, 14, 2

fviz_cluster(FIRM.WORTH, data = FIRM.WORTH.DF) ##Visual
```
##A high PE Ratio can be seen as good or bad, depending. Since these are established firms rather than a startup, we want a lower PE Ratio indicating that we are not overpaying for the value of the stock. A higher market capitalization is always better
##The firm with the best stock value (you should invest) is 13 - Johnson & Johnson (in my opinion)
##Options in cluster 1 have the best stock value

##Are other attributes are valuable to look at, but the 4 I have chosen to concentrate on will lead to the best odds of high performance if stock is purchased. Looking into the other attributes may cloud the waters.

##If only one stock can be purchased than the best option would be Pfizer Inc who is located in high performing ares in their clusters of both segmentations.
