---
title: "DeSimone_MS64060_Assignment 3"
author: "Heather DeSimone"
date: "3/5/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

##First I have loaded in my data frame and called a summary of the information.
```{r}
DF=read.csv("C:/Users/hdesi/Desktop/MBA/Machine Learning/UniversalBank2.csv")
library(dplyr)
library(caret)
library(class)
library(ISLR)
DF <- DF %>% relocate(Personal.Loan, .after = CreditCard)
summary(DF)
```
##I have converted a few attributes over to factors - these attributes classify a yes (1) or no (0) response.I have called a summary to check my work.
```{r}
DF$Personal.Loan=as.factor(DF$Personal.Loan)
DF$Securities.Account=as.factor(DF$Securities.Account)
DF$CD.Account=as.factor(DF$CD.Account)
DF$Online=as.factor(DF$Online)
DF$CreditCard=as.factor(DF$CreditCard)
summary(DF)
```
##Question A
##I will now separate my data into training and validating sets - training = 60% and validation = 40%.
##I have also created my pivot table.
```{r}
Train_Index = createDataPartition(DF$Personal.Loan,p=0.6, list=FALSE)
Train.df=DF[Train_Index,]
Validation.df=DF[-Train_Index,] 

mytable <- xtabs(~ CreditCard+Online+Personal.Loan, data=Train.df)
ftable(mytable) 
```
##Question B
##The probability that a customer will accept a loan offer based condionally that they have a credit card and online account is roughly 10% (.0996)


##Question C
##Creating my 2 new pivot tables
```{r}
table(Personal.Loan=Train.df$Personal.Loan, Online=Train.df$Online)
table(Personal.Loan=Train.df$Personal.Loan, CreditCard=Train.df$CreditCard)
```
##Question D

##i. P(CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors)
## Answer is 92/(196+92) = .319 = 32%

##ii. P(Online = 1 | Loan = 1)  (the proportion of Online users among the loan acceptors)
## Answer is 172/(116+172) = .597 = 60%

##iii. P(Loan = 1) (the proportion of loan acceptors)  
## Answer is (196+92)/(1917+795+196+92) = 288/3000 = .096 = 10%

##iv. P(CC = 1 | Loan = 0)  
##Answer is 795/(795+1917) = .293 = 29%

##v. P(Online = 1 | Loan = 0) 
## Answer is 1594/(1118+1594) = .587 = 59%

##vi. P(Loan = 0) 
## Answer is (1917+795)/(1917+795+196+92) = .904 = 90%

##Question E
##P(Loan = 1 | CC = 1, Online = 1).
##P(Loan = 1) = .319*.597 = .19

##Question F
##The pivot table is more accurate because there are more variables used in the prediction. The Naive Bayes assumes that each prediction is independent from each variable.

##Question G
##Running Naive Bayes on the data
```{r}
library(e1071)
nb.model<-naiveBayes (Personal.Loan~CreditCard+Online, data=Train.df)
To_Predict=data.frame(CreditCard='1', Online='1')
predict(nb.model,To_Predict,type='raw')
```
##The above running of the naive bayes on my data is very close to the prediction I made in Question B. Question E has a very different answer than B and G. I would conclude that 10% is the correct prediction.

